---
title: "CSC8631 Report"
author: "Morgan Frodsham"
date: "08/11/2021"
output: pdf_document
citation_package: natbib
bibliography: "references.bib"
biblio-style: "apalike"
link-citations: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir= normalizePath('..'))
```

```{r ProjectTemplate, include = FALSE}
library(ProjectTemplate)
load.project()
```

# FutureLearn Learning Analytics for Newcastle University

## Business understanding

### Background

This report details the process and findings of exploratory data analysis for Newcastle University's FutureLearn course called "Cyber Security: Safety at Home, Online, in Life". Data analysis of educational platforms, such as FutureLearn, is a key part of learning analytics, which is defined as *"the measurement, collection, analysis and reporting of data about learners and their contexts, for purposes of understanding and optimising learning and the environments in which it occurs."* @Learning-Analytics-Conference Learning analytics has the potential to help educational institutions undertake data-driven decision-making @EDUCAUSE that better support students through their educational journey @Bricks.

### Business objectives and success criteria

FutureLearn is a massive open online course (MOOC) provider. It promotes itself as *"a powerful new way to learn online"* @FurtureLearn, designed with the principles of effective learning. Learning effectiveness can be indicated by student engagement, their success rate, and completion time @E-learning. A core objective of FutureLearn is to improve learning effectiveness.

The objective of this data analysis is to better understand the learning effectiveness of Newcastle University's Future Learn Cyber Security course. The process and findings outlined by this report explore the question: "Do learners have a higher chance of success the more data they share with FutureLearn?" Here, "a higher chance of success" means that learners are more likely to complete the course. The data shared by learners refers to the optional data requested by FutureLearn; specifically, the provision of personal identifiers like demographic information. The provision of this data is taken to be an indicator of student engagement with the course. For this analysis the business success criteria is to give useful insights into the relationship between learner engagement and learner success rate.

(objectives that were considered and rejected, and rational for why?)

### Inventory of resources

Newcastle University has provided x number of data sets for exploration.

Tidyverse for R is the library used for the analysis, ggplot2 for R is the library used for visualisations, ProjectTemplate for R is used to improve the reproducibility of this analysis, Git is used for version control, and all written documentation is created through RMarkdown for R.

### Requirements, assumptions and constraints

This project, as well as the process and findings outlined in this report, is limited to the data analysis pipeline. (assumptions made about the nature of the problem, constraints imposed on the project (technical suite)?) corellation, confounding variables

### Risks and contingencies

Risk re un-reproducibility, version control, etc. - check data flow lecture) The technical suite used aims to apply good programming practice, and reduce common risks that might delay the project or cause it to fail.

### Terminology

### Costs and benefits

This section describes the costs of the project and predicted business benefits if the project is successful (e.g., return on investment). Other less tangible benefits (e.g., customer satisfaction) should also be highlighted.

### Data mining goals and success criteria

The initial investigation outlined by this report explores the question: "Do learners have a higher chance of success the more data they share with FutureLearn?" To understand this, the data mining aims to determine:

1.  What is the success rate of learners? The data mining success criteria is the percentage of all learners who complete the course.

2.  How many learners provide any demographic data? The data mining success criteria is the percentage of all learners who provide optional demographic data such as gender, age range or employment status.

3.  What is the success rate of learners who provide demographic data? The data mining success criteria is the percentage of learners who complete the course and provide demographic data.

4.  What is the success rate of learners who do not provide demographic data? The data mining success criteria is the percentage of learners who complete the course and do not provide any demographic data.

5.  Is there any correlation between success and the provision of demographic data? The data mining success criteria is the difference between (and standard deviation of?) each of the aforementioned learner success rates.

### Project plan

The deadline for this project is 16:30, 3 December 2021. (This section lists the stages to be executed in the project, together with their duration, resources required, inputs, outputs, and dependencies. Where possible, it should make explicit the large-scale iterations in the data mining process---for example, repetitions of the modeling and evaluation phases)

### Initial assessment of tools and techniques

(methodology, This section gives an initial view of what tools and techniques are likely to be used and how. It describes the requirements for tools and techniques, lists available tools and techniques, and matches them to requirements.)

## Data understanding

### Initial data collection

Newcastle University's FutureLearn cyber security course has run seven times. The first run provided six different databases, the second run provided seven, and the others provided eight. There are 53 different databases for potential exploration.

There have been no problems in data acquisition or extraction.

### Data description

The first run of Newcastle University's FutureLearn cyber security course provides the following databases:
- archetype survey responses (id, learner_id, responded_at, archetype); 
- enrolments (learner_id, enrolled_at,	unenrolled_at	role,	fully_participated_at,	purchased_statement_at	gender,	country	age_range,	highest_education_level,	employment_status,	employment_area	detected_country); 
- leaving survey responses (id, learner_id,	left_at	leaving_reason, last_completed_step_at,	last_completed_step,	last_completed_week_number,	last_completed_step_number); 
- question responses (learner_id, quiz_question,	question_type,	week_number,	step_number,	question_number,	response,	cloze_response,	submitted_at,	correct); 
- step activity (learner_id	step,	week_number	step_number,	first_visited_at, last_completed_at); and
- weekly sentiment survey (id	responded_at,	week_number,	experience_rating,	reason).

The second run provides the same databases as well as a databases specifying the course team members (id,	first_name,	last_name,	team_role,	user_role). 

The other five runs provide the same databases as the second run as well as a database on the course's video stats (step_position,	title	video_duration,	total_views,	total_downloads,	total_caption_views,	total_transcript_views,	viewed_hd,	viewed_five_percent,	viewed_ten_percent,	viewed_twentyfive_percent,	viewed_fifty_percent,	viewed_seventyfive_percent,	viewed_ninetyfive_percent,	viewed_onehundred_percent,	console_device_percentage,	desktop_device_percentage,	mobile_device_percentage,	tv_device_percentage,	tablet_device_percentage,	unknown_device_percentage,	europe_views_percentage,	oceania_views_percentage,	asia_views_percentage,	north_america_views_percentage,	south_america_views_percentage,	africa_views_percentage,	antarctica_views_percentage).

### Data exploration



The exploratory analysis outlined by this report is conducted on the enrolments and archetypes data sets for all seven runs. The data sets for all seven runs have been provided by Newcastle University for the analysis in this report.

There are 37296 learners in the seven runs of Newcastle University's FutureLearn Cyber Security course, and each learner has a unique learner_id. In the enrolments and archetypes data sets there are 13 and four columns of data, respectively. There are some missing values in the enrolments data set (what about archetypes?), but otherwise the data quality is good.


To understand the FutureLearn students' success rate (course completion) for Newcastle University's Cyber Security course, a variable was created from filtering the students by those who had data under fully_participated_at.

(The success criteria of this project will include course completion as success rate, provision of personal identifiers such as demographic information as student engagement, and completion time?. Data analysis will be focused on the enrollments data, and use CRISP-DM as the methodology.) To become familiar with the data, the enrollments data sets for all 7 runs were combined into one full dataset.


### Data quality 

(Completeness/accuracy - not completely complete)
p
The data collected by the first run of the course is relatively sparse compared to the seventh run of the course. In runs three and four more data is being collected; this could be because there are fewer optional sections (such as the archetype or leaving survey), but it is unknown why this change has occurred.

## Data preparation 

(a description of the dataset after pre-processing and the process it was produced - basically why and how of munge stuff)

```{r cars}
summary(cars)
```



```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.



\newpage

# Bibliography
