---
title: "CSC8631 Report"
author: "Morgan Frodsham"
date: "08/11/2021"
output: pdf_document
citation_package: natbib
bibliography: "references.bib"
biblio-style: "apalike"
link-citations: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir= normalizePath('..'))
```

```{r ProjectTemplate, include = FALSE}
library(ProjectTemplate)
load.project()
```

# Learning Analytics for Cyber Security: Safety at Home, Online, in Life

## Objectives

This report details the process and findings of exploratory data analysis for Newcastle University's FutureLearn course called "Cyber Security: Safety at Home, Online, in Life". Data analysis of educational platforms, such as FutureLearn, is a key part of learning analytics, which is defined as *"the measurement, collection, analysis and reporting of data about learners and their contexts, for purposes of understanding and optimising learning and the environments in which it occurs."* @R-base (Add reference - 1st International Conference on Learning Analytics and Knowledge, Banff, Alberta, February 27--March 1, 2011, <https://tekri.athabascau.ca/analytics/>.) Learning analytics has the potential to help educational institutions undertake data-driven decision-making (reference - <https://er.educause.edu/articles/2011/9/penetrating-the-fog-analytics-in-learning-and-education>) that better support students through their educational journey (reference - from bricks to clicks)

FutureLearn is a massive open online course (MOOC) provider (reference - <https://www.futurelearn.com/info/blog/what-is-a-mooc-futurelearn>). It promotes itself as *"a powerful new way to learn online"* (Reference - <https://www.futurelearn.com/using-futurelearn/why-it-works>) that is designed with the principles of effective learning. Learning effectiveness can be indicated by student engagement, their success rate, and completion time (reference - <https://reader.elsevier.com/reader/sd/pii/S0747563218302590?token=A32E75AFAA1095275B9E0B7D3FA82CF29E5B168C4C92AE4B3D48A26C6456C9A142B81B117C93C71602CF948FB87D8F31&originRegion=eu-west-1&originCreation=20211120115140>). The objective of this data analysis is to better understand the learning effectiveness of Newcastle University's Future Learn Cyber Security course.

(The success criteria of this project will include course completion as success rate, provision of personal identifiers such as demographic information as student engagement, and completion time?. Data analysis will be focused on the enrollments data, and use CRISP-DM as the methodology.) To become familiar with the data, the enrollments data sets for all 7 runs were combined into one full dataset.

## Course completion

To understand the FutureLearn students' success rate (course completion) for Newcastle University's Cyber Security course, a variable was created from filtering the students by those who had data under fully_participated_at.

```{r cars}
summary(cars)
```

This showed that 2154 of 37296 students completed the course. \[How can I tell if all the student IDs are unique? Should I do that? How do I check that all the IDs are of the role "learner"?\]

## Gender

Of those who had completed the course, gender identty.

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Country

## Age Range

## Highest Education Leave

## Employment Status

## Employmenmt Area

## Course Completed and Amount of Information Provided

## Correlation between different types of information and course completion

\newpage
# Bibliography
